The \emph{FThresholds} package for \emph{Macaulay2}

### 2019-06-06 Fsing/FThresholdsPaper/FThresholds.tex
minor updates to paper

\author[]{Emily E.\ Witt}
\address{Department of Mathematics, University of Kansas, Lawrence, KS~66045, USA}
\email{witt@ku.edu}
\thanks{E.E.~Witt was partially supported by NSF DMS \#1623035.}
\date{\today}

\begin{abstract}
   This note describes the functionality implemented in the \emph{Macaulay2} package \emph{FThresholds}.
   This package is designed to compute and estimate $F$-pure thresholds, more general $F$-thresholds, and related numerical invariants arising in the study of singularities in positive characteristic commutative algebra.
\end{abstract}

@@

o6 : QQ
\end{MyVerbatim}
}
\medskip
The numbers $5/7, 40/49, 285/343$ then approximate the $F$-pure threshold, which in this case happens to equal $5/6$.  By default, if the function $f$ has a special form (is diagonal or binomial), then the function does not check whether $f^a$ is in  $\langle x_1^{p^e}, \dots, x_n^{p^e} \rangle$, but instead uses general formulas for $\nu$s, see \cite{HernandezFPureThresholdOfBinomial,HernandezFInvariantsOfDiagonalHyp}.

@@

The command {\tt isFPT(t, f)} checks whether $t$ is the $F$-pure threshold of $f$.  In general this does this by computing two test ideals:
\[
\tau(R, f^t) \text{ and } \tau(R, f^{t-\epsilon}).
\]
It follows that $\fpt(f) = t$ if and only if $\tau(R, f^t) \neq R$ and $\tau(R, f^{t-\epsilon}) = R$.  The function computes these two values by using the {\tt testIdeal} and {\tt FPureModule} commands from the {\tt TestIdeals} package.  Note that in a strongly $F$-regular Gorenstein ambient ring, {\tt FPureModule(t, f)} computes exactly $\tau(R, f^{t-\epsilon})$.  These commands work more generally, but they \emph{only} currently work when $R$ is $\bQ$-Gorenstein with index not divisible by the characteristic $p > 0$.  A more general version of this command is {\tt compareFPT(t, f)} which returns $-1$ if $t < \fpt(f)$, $0$ if $t = \fpt(f)$ and $1$ if $t > \fpt(f)$.

>>>>

\author[]{Emily E.\ Witt}
\address{Department of Mathematics, University of Kansas, Lawrence, KS~66045, USA}
\email{witt@ku.edu}
\thanks{E.~E.~Witt was partially supported by NSF DMS \#1623035.}
\date{\today}

\begin{abstract}
   This note describes the functionality implemented in the \emph{Macaulay2} package \emph{FThresholds}.
   This package is designed to compute and estimate $F$-pure thresholds, more general $F$-thresholds, and related numerical invariants arising in the study of singularities in prime characteristic commutative algebra.
\end{abstract}

@@

o6 : QQ
\end{MyVerbatim}
}
\medskip
The numbers $5/7, 40/49, 285/343$ then approximate the $F$-pure threshold, which in this case happens to equal $5/6$.  By default, if the function $f$ has a special form (is a diagonal polynomial or a binomial), then the function does not check whether $f^a$ is in  $\langle x_1^{p^e}, \dots, x_n^{p^e} \rangle$, but instead uses general formulas for $\nu$s, see \cite{HernandezFPureThresholdOfBinomial,HernandezFInvariantsOfDiagonalHyp}.

@@

Given a rational number $t$ and  an element $f$ of a $\mathbb{Q}$-Gorenstein ring $R$ of characteristic $p>0$, 
the command {\tt isFPT(t, f)} checks whether $t$ is the $F$-pure threshold of $f$.  In general, this is done by computing the \emph{test ideals}
$\tau(R, f^t)$ and $\tau(R, f^{t-\epsilon})$; 
the $F$-pure threshold equals $t$ if and only if $\tau(R, f^{t-\epsilon}) = R$, but  $\tau(R, f^t)$ is proper.  The function computes these two values by using the {\tt testIdeal} and {\tt FPureModule} commands from the {\tt TestIdeals} package.  Note that in a strongly $F$-regular Gorenstein ambient ring, {\tt FPureModule(t, f)} computes exactly $\tau(R, f^{t-\epsilon})$.  These commands work more generally, but they \emph{only} currently work when $R$ is $\bQ$-Gorenstein with index not divisible by the characteristic $p > 0$.  A more general version of this command is {\tt compareFPT(t, f)} which returns $-1$ if $t < \fpt(f)$, $0$ if $t = \fpt(f)$ and $1$ if $t > \fpt(f)$.
